---
title: "Example 9.6"
author: "Monnie McGee"
output: 
  html_document:
    keep_md: true
---


The purpose of this case study is to use R to explore different on-line job postings for different positions in ''data science''. The assumption is that we can use the job descriptions that employers post for data scientists to define what a data scientist does (and thereby arrive at a definition of data science as a discipline).

### Deliverable
In using data science to define data science, we need to keep in mind reproducibility. To this end, your project markdown code should be committed and pushed to GitHub. The deliverable is a link to the GitHub repository where the markdown code resides. 

### The Code
 We will make a separate function for each step, which should make the functions easier to read, test, maintain, and adjust as the format of the web pages changes. The function `cy.getFreeFormWords()` below fetches the lists of free-form text in the HTML document. The function then decomposes the text into the words in each element, using spaces and punctuation characters to separate them. This is done by calling the `asWords()` function. One of the arguements for `asWords()` is a list of "stop words", which are small words that are present in a large number of English sentences. We don't want to include these words in our list of post words. Finally, a call to `removeStopWords()` removes all stop words from the post, so that we have only the words that carry meaning for the job seeker (well, almost).

```{r Load libraries}
library(XML)
library(RCurl)
```

```{r getFreeForm}
StopWords = readLines("http://jmlr.csail.mit.edu/papers/volume5/lewis04a/a11-smart-stop-list/english.stop")

asWords = function(txt, stopWords = StopWords, stem = FALSE)
{
  words = unlist(strsplit(txt, '[[:space:]!.,;#:()/"]+'))
  words = words[words != ""]
  if(stem && require(Rlibstemmer))
     words = wordStem(words)
  i = tolower(words) %in% tolower(stopWords)
  words[!i]
}

removeStopWords = function(x, stopWords = StopWords) 
     {
         if(is.character(x))
             setdiff(x, stopWords)
         else if(is.list(x))
             lapply(x, removeStopWords, stopWords)
         else
             x
     }

cy.getFreeFormWords = function(doc, stopWords = StopWords)
     {
         nodes = getNodeSet(doc, "//div[@class='job-details']/
                                 div[@data-section]")
         if(length(nodes) == 0) 
             nodes = getNodeSet(doc, "//div[@class='job-details']//p")
         
         if(length(nodes) == 0) 
             warning("did not find any nodes for the free form text in ",
                     docName(doc))
         
         words = lapply(nodes,
                        function(x)
                            strsplit(xmlValue(x), 
                                     "[[:space:][:punct:]]+"))
         
         removeStopWords(words, stopWords)
     }

```

### Question 1: Implement the following functions. Use the code we explored to extract the date posted, skill sets and salary and location information from the parsed HTML document.

```{r Question1}
cy.getSkillList = function(doc)
{
  lis = getNodeSet(doc, "//div[@class = 'skills-section']//
                         li[@class = 'skill-item']//
                         span[@class = 'skill-name']")

  sapply(lis, xmlValue)
}

cy.getDatePosted = function(doc)
  { xmlValue(getNodeSet(doc, 
                     "//div[@class = 'job-details']//
                        div[@class='posted']/
                        span/following-sibling::text()")[[1]],
    trim = TRUE) 
}

cy.getLocationSalary = function(doc)
{
  ans = xpathSApply(doc, "//div[@class = 'job-info-main'][1]/div", xmlValue)
  names(ans) = c("location", "salary")
  ans
}

# cy.getSkillList(cydoc)
# cy.getLocationSalary(cydoc)
```

The function `cy.ReadPost()` given below reads each job post. This function implements three other functions: `cy.getFreeFormWords()`, `cy.getSkillList()`, and `cy.getLocationSalary()`.

```{r cy.readPost}
cy.readPost = function(u, stopWords = StopWords, doc = htmlParse(u))
  {
    ans = list(words = cy.getFreeFormWords(doc, stopWords),
         datePosted = cy.getDatePosted(doc),
         skills = cy.getSkillList(doc))
    o = cy.getLocationSalary(doc)
    ans[names(o)] = o
    ans
}
# cyFuns = list(readPost = function(u, stopWords = StopWords, doc=htmlParse(u)))
```
### Reading posts programmatically
The function `cy.ReadPost()` allows us to read a single post from CyberCoders.com in a very general format. All we need is the URL for the post. Now, let's see about obtaining the URLs using a computer program.

```{r GetPosts}
# Obtain URLs for job posts
txt = getForm("http://www.cybercoders.com/search/", searchterms = '"Data Scientist"',
              searchlocation = "",  newsearch = "true", sorttype = "")
# Parse the links
doc = htmlParse(txt, asText = TRUE)
links = getNodeSet(doc, "//div[@class = 'job-title']/a/@href")
# Save the links in the vector joblinks
joblinks <- getRelativeURL(as.character(links), "http://www.cybercoders.com/search/")
# Read the posts
# posts <- lapply(joblinks,cy.readPost)

cy.getPostLinks = function(doc, baseURL = "http://www.cybercoders.com/search/") 
  {
    if(is.character(doc)) doc = htmlParse(doc)
    links = getNodeSet(doc, "//div[@class = 'job-title']/a/@href") 
    getRelativeURL(as.character(links), baseURL)
}

cy.readPagePosts = function(doc, links = cy.getPostLinks(doc, baseURL),
baseURL = "http://www.cybercoders.com/search/")
  {
    if(is.character(doc)) doc = htmlParse(doc)
    lapply(links, cy.readPost)
 }

## Testing the function with the parsed version of the first page of results in object doc
posts = cy.readPagePosts(doc)
sapply(posts,`[[`, "salary")
summary(sapply(posts, function(x) length(unlist(x$words))))
```

**Question:** Test the `cy.getFreeFromWords()` function on several different posts.

The following code chunk pulls it all together. The function `cy.getNextPageLink()` retrieves each page from CyberCoders and calls the other functions to parse each post in order to obtain information such as salary, skills, and location.

```{r Next Page of Results}
# Test of concept
# getNodeSet(doc, "//a[@rel='next']/@href")[[1]]
## A function to get all pages
cy.getNextPageLink = function(doc, baseURL = docName(doc))
{
  if(is.na(baseURL))
     baseURL = "http://www.cybercoders.com/"
  link = getNodeSet(doc, "//li[@class = 'lnk-next pager-item ']/a/@href")
  if(length(link) == 0)
    return(character())
    link2 <- gsub("./", "search/",link[[1]])
 getRelativeURL(link2, baseURL)
}

# Test the above function
# tmp = cy.getNextPageLink(doc, "http://www.cybercoders.com")
```

Now we have all we need to retrieve all job posts on Cyber Coders for a given search query. The following function puts it all together into a function that we can call with a search string for a job of interest. The function submits the initial query and then reads the posts from each result page.
```{r cyberCoders}
cyberCoders =
function(query)
{
   txt = getForm("http://www.cybercoders.com/search/",
                  searchterms = query,  searchlocation = "",
                  newsearch = "true",  sorttype = "")
   doc = htmlParse(txt)

   posts = list()
   while(TRUE) {
       posts = c(posts, cy.readPagePosts(doc))
       nextPage = cy.getNextPageLink(doc)
       if(length(nextPage) == 0)
          break

       nextPage = getURLContent(nextPage)
       doc = htmlParse(nextPage, asText = TRUE)
   }
   invisible(posts)
}
```

The function cyberCoders is called below with the skill "Data Scientist". Then, we sort the skills and obtain all skills that are mentioned more than twice in the list.

```{r Get Skills, cache=TRUE}
dataSciPosts = cyberCoders("Data Scientist")
tt = sort(table(unlist(lapply(dataSciPosts, `[[`, "skills"))),
           decreasing = TRUE)
tt[tt >= 2]
```


Reference:
Code taken from Nolan, D. and Temple Lang. Data Science in R: A Case Studies Approach to Computational Reasoning and Problem Solving. CRC Press, 04/2015. VitalBook file.
